#Lab_04 {lab_04}

rob wells, 8/14/2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)

# Load janitor here

library(janitor)


```

### Load data

In this data set, each row represents a city or town in Maryland with the 2010 and 2020 figures for: 

-  Total population 
-  White only population
-  Black or African American only population
-  Hispanic only population 
-  Native American only population
-  Asian only population
-  Pacific Islander (including native Hawiian) only population

Background on this data from [Social Explorer](https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001) 
```{r}
# Write code to load the Maryland race by city, place for 2010 - 2020 and call it md_race.
# The data is called lab_03.csv, and is in the same folder as lab_03.qmd (this data notebook)
md_race<- rio::import("lab_04.csv")
```

## Answer questions

Q1: Name the five cities that had the largest Hispanic population as a percentage of the entire population in 2020 and call it top5_hispanic. Produce a top 5 table for each, selecting the place, 2020 Hispanic, 2020 total and your new calculation.

A1:
```{r}

# Write your code in this codeblock.


top5_hispanic <-md_race %>% 
  select(place, x2020_hispanic, x2020_total) %>% 
   mutate(Percent_total = (x2020_hispanic/x2020_total)) %>% 
  slice_max(Percent_total,n=5)
 #head(large_hispanic2020,5)  

```

Q2: Compare the top 10 cities with the largest Black population as a percentage of the overall population in 2010 and 2020. Filter for places with more than 20,000 population in 2010. Call your results pct_black. Compare results for 2020 vs 2010. What is different? What surprised you? 

A2:

```{r}
#write your code here
pct_black <-md_race %>% 
  select(place, x2010_black, x2020_black,x2010_total, x2020_total) %>% 
  (x2010_black > 20000) %>% 
slice_max(x2020_black,n=10) %>% 
 mutate(Percent_total = (x2020_total/x2010_total)) 
#arrange(desc(x2020_black))
#pct_black
#your selet lines must have all the varibles, you cannot filter anything unless it has been sleected. It won't exisit if its not already in the datatset. 
# It was interesting to see what cities had high increases in the black population fom 2010 to 2020 such as Waldorf, Bowie, and Clinton. very few had sharp decreases, however some did remain close to to constant number such as Woodlawn CDP 
```

Q3: Find the 15 communities that had the great gains in Hispanic population from 2010 to 2020 as measured by percentage change. Also calculate percentage change for the community's overall growth during the same period. Filter your results for communities with more than 10,000 people as of 2020. Your results will be hisp_growth. Look at your results, craft a one-paragraph story pitch for your editor. 

A3:

```{r}
hisp_growth <-md_race %>% 
  select(place,x2020_hispanic,x2010_hispanic,x2020_total) %>% 
  mutate(Percent_change = (x2020_hispanic-x2010_hispanic)/x2010_hispanic) %>% 
   slice_max(x2020_hispanic,n=15) %>% 
filter(x2020_total > 10000) 


#After reveiwing the data that displayed the 15 communities with the greatest gains form 2010-2020, i found that Baltimore city was at the top of that list. Wheaton and Germantown followed right along with silver Spring or Aspen Hill. While this is not too surprising because cities such as Baltimore  tend to attract more minorities, it would be fascinating learn the various resoruces that are offered in these areas that may be attracting a larger Hispanic population. Germantown is one of the most diverse cities in Montgomery county. I also think it would be intresting to learn why Annapplis and Aldelphi seem to lose more Hispanics than any of the other cities. 


```
	demo(colo
	
	
	summarize(x2010_hispanic, newColName = someFunction())
	