#Lab_03 {lab_03}

rob wells, 8/14/2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
install.packages("tidyverse")
install.packages("janitor")
# Load janitor here


```

### Load data

In this data set, each row represents a city or town in Maryland with the 2010 and 2020 figures for: 

-  Total population 
-  White only population
-  Black or African American only population
-  Hispanic only population 
-  Native American only population
-  Asian only population
-  Pacific Islander (including native Hawiian) only population

Background on this data from [Social Explorer](https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001) 

library(tidyverse)
library(rio)
library(janitor)
```{r}
# Write code to load the Maryland race by city, place for 2010 - 2020 and call it md_race.
# The data is called , and is in the same folder as lab_03.qmd (this data notebook)
 md_race<- rio::import("lab_03.csv")
 
# your answer here

```


## Answer questions

Q1: Which five cities have the largest overall Hispanic population in 2020? Name the five cities that had the largest Hispanic population in 2020. 

A1:
```{r}

# Write your code in this codeblock.
largest5_hispanic2020 <-md_race %>% 
  select(place, x2020_hispanic) %>% 
slice_max(x2020_hispanic,n=5)

largest5_hispanic2020
# your answer here
#The five cities inlucde Baltimore city, wheaton, germantown, silver spring, aspen hill 
```

Q2: Find the cities with the largest Black population and arrange the results in descending order. 

A2:

```{r}
#write your code here
large_blackpop <-md_race %>% 
  select(place, x2020_black) %>% 
#slice_max(x2020_black,n=5)
arrange(desc(x2020_black))
large_blackpop

# your answer here
#Baltimore city
```

Q3: Create a table called top_hispanic that shows the top 15 communities with the largest Hispanic population in 2020, and also display the 2010 Hispanic population and the total community population for 2020. Look at your results - what stands out to you? Craft a one-paragraph story pitch for your editor. 

A3:

```{r}
large_hispanic2020 <-md_race %>% 
  select(place,x2020_hispanic,x2010_hispanic,x2020_total) %>% 
  #arrange(desc(x2020_hispanic))
   slice_max(x2020_hispanic,n=15)
large_hispanic2020
 #head(large_hispanic2020,15)  

# your answer here
#After looking at the results from this data that pulled the population of Hispanic communities  from 2010-200 throughout Maryland, I found that the Hispanic population grew in B altimore form 2010 to 2020 In Baltimore city, Wheaton, Germantown, Aspen Hill, and Gaithurburg. Ultimately, the Hispanic population grew throughout Majority of the cities in Montgomery County. I think the an interesting angle would be to look into. the polcies that may have encourged this influx and why they to be flocking in high numbers to montogmery county, a county with one of the highest median incomes
```
