---
title: "Data Anaylsis Project"
output: html_document
date: "2022-11-21"
---


# Ben Reitman and Grace Kpetemey
#WELLS COMMENT. YOU AND GRACE HANDED IN SEPARATE VERSIONS, WHICH WERE SIMILAR IN MANY WAYS. If you are doing a group project, hand in one version.  Going forward, hand in one assignment for the both of you, like I discussed in class.
#Because Ben's version had charts, I provided detailed comments after question #1 on his version. Grace, work with Ben's version for the next stage of this project

#I posted this comment on your first draft notebook. It's important and I want you to address it immediately:
#I want you both to do the following before the Tuesday class:
#1. Conduct a search on nces.ed.gov for the total population of school students. You need that for your denominator to calculate the rate of suspensions. I did a very quick Google search and found the data. 



# Our project is based around analyzing data regarding students and school discipline when students get expelled or suspended. Our data set comes from the National Center for Education Statistics and breaks down discipline categories by race, gender, race AND gender, and also has a separate category for tracking discipline for students with disabilities. In order to get the data set to be able to used in R, I had to manually clean some of the data and give each specific variable a different name. For example, every variable with "M_" in front is for male, "F_" is female, "MD_" is for male students with disabilities and "FD" is for female students with disabilites. For this project, we are mainly focusing on all_suspensions, and all_expulsions. 

#WELLS COMMENT: REMOVE THIS - UNNECESSARY
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Wells comment: Remove scientific notation: options(scipen = 999)

options(scipen = 100, digits = 4)
options(scipen = 999)
library(tidyverse)
library(janitor)
```
#WELLS COMMENT: SEPARATE CHUNKS WITH TWO LINES OR MORE


```{r}
Education_Stats_CLEAN <- read_csv("Education Stats - Copy of Digest 2018 Table 233.27.csv") %>% 
  clean_names()

```


#1: What is the rate for total suspensions and expulsions combined for all students versus students with disabilities. 

```{r}
suspensions_expulsions<-Education_stats_CLEAN %>% 
  select(disability_status_sex_and_race_ethnicity, all_suspensions, all_expulsions) %>% 
  filter((disability_status_sex_and_race_ethnicity == "All_Students") | (disability_status_sex_and_race_ethnicity == "Disabled_students")) %>% 
  mutate(combined = all_suspensions + all_expulsions)


```



```{r}

ggplot(suspensions_expulsions, aes(all_suspensions, disability_status_sex_and_race_ethnicity, fill = all_suspensions)) +
geom_col(position = "dodge")+
labs(
  title = "Suspensions for Disabled Students comapred to All")


```



```{r}
ggplot(suspensions_expulsions, aes(all_expulsions, disability_status_sex_and_race_ethnicity, fill = all_expulsions)) +
geom_col(position = "dodge")+
labs(
  title = "Expulsions for Disabled Students comapred to All")

```

# Both of these charts show that the rate of both suspensions and expulsions for disabled students to all students is practically 
#SPELLING# exactlt the same.






# Our table shows that "total" (the total number of suspensions) and "all_expulsions" added up gives us the combined number of suspensions and expulsions for all students, and students with disabilites.




#2:  What is the rate of total suspensions to total expulsions? For every expulsion, how many student suspensions are there?


```{r}
Education_stats_CLEAN %>% 
  select(disability_status_sex_and_race_ethnicity, all_suspensions, all_expulsions) %>% 
  filter((disability_status_sex_and_race_ethnicity == "All_Students") | (disability_status_sex_and_race_ethnicity == "Disabled_students")) %>% 
  mutate(Suspension_rate = all_suspensions/all_expulsions)

# create ratio of suspension and expulsion rate of disabled to all students
```



```{r}
Disabled_suspensions_ratio = (644605/2635743)*100
Disabled_all_expulsions = (26852/111215)*100
```
This shows us the ratio of all of the all suspensions for kids with disabilities to all students as well as stats for all expulsions. It shows us that these rates are the same and that it feels like this not by accident. There are exactly at 24 percent.
#WELLS COMMENT: NOT EXACTLY THE SAME. THE suspensions are 24.5 and the expulsions are 24.1
#The broad issue involves a comparison to the entire universe of students. This is why I asked you to find the total population of students to make this comparison.  Conduct a search on nces.ed.gov for the total population of school students. You need that for your denominator to calculate the rate of suspensions. I did a very quick Google search and found the data. 




This table gives us a result that might be a bit surprising. This shows that for all students in schools, as well as students with disabilities, they both have around the same rate of suspension to expulsion rate despite students with disabilities have a lower total number of students. For around every 24 students that get suspended, one #SPELLING# xtudent gets expelled for both the overall student population and the disabled student population.


#3:  How many suspensions or expulsions do minority children face in comparison to their white counterparts?

#WELLS COMMENT: THIS table is good but to tell the entire story you need to look at the entire universe of students. This is why I asked you to find the total population of students to make this comparison.

```{r}
by_race <- Education_stats_CLEAN %>% 
  select(disability_status_sex_and_race_ethnicity, all_suspensions, all_expulsions) %>% 
  filter((disability_status_sex_and_race_ethnicity == "All_White") |(disability_status_sex_and_race_ethnicity == "All_Black") | (disability_status_sex_and_race_ethnicity == "All_Hispanic") | (disability_status_sex_and_race_ethnicity == "All_Asian") | (disability_status_sex_and_race_ethnicity == "All_Pacific Islander") | (disability_status_sex_and_race_ethnicity == "All_American Indian/Alaska Native") | (disability_status_sex_and_race_ethnicity == "All_Two or more races")) %>% 
mutate(discipline_by_race = all_suspensions + all_expulsions)
```
This table allows us to look at school discipline by race and that despite being a minority, it is very clear that black students get discipined the most in schools with about 18,000 more cases of discipline than white students.

#Graphic 3
#wells comment: sort this table according to the values in descending order

```{r}

ggplot(by_race, aes(discipline_by_race, disability_status_sex_and_race_ethnicity, fill = discipline_by_race)) +
geom_col(position = "dodge")+
labs(
  title = "Discipline in Schools by Race")

``` 


#4: What is the difference between of white males and black males who #SPELLING# recieve corproal pubishment? What is that same difference between black females and white females?

```{r}
male <- Education_stats_CLEAN %>% 
  select(disability_status_sex_and_race_ethnicity,corporal_punishment) %>% 
   filter((disability_status_sex_and_race_ethnicity == "M_White") |(disability_status_sex_and_race_ethnicity == "M_Black") | (disability_status_sex_and_race_ethnicity == "F_White") | (disability_status_sex_and_race_ethnicity == "F_Black")) 

#print(sapply(race_cp, class))

#WELLS COMMENT: THIS BELOW SEEMS UNNECESSARY. DONT KNOW WHY YOU DID IT BECAUSE THE CODE ABOVE DID THE SAME THING
race_corp=data.frame(disability_sex_race=c("M_White","M_Black","F_White","F_Black"),corporal_punishment=c(44319, 29840, 8750, 10183))
``` 

#Graphic 4
#WELLS COMMENT - AGAIN, THIS IS WHERE THE TOTAL POPULATION FIGURES WILL PROVIDE CONTEXT
```{r}
race_corp %>% 
  ggplot() +
  geom_bar(aes(x=reorder(disability_sex_race, corporal_punishment), weight=corporal_punishment), fill = "blue") +
  coord_flip() + 
  labs(
    title="differences of corpral punishment by race and gender ",
    x = "",
    y = "amount"
  )

#WELLS COMMENT: ERROR, THE CHUNK NEEDS TO HAVE THREE APOSTOPHES TO CLOSE IT
```

#WELLS COMMENT - Please write a brief summary of what you found with this chunk


#5: What is the percetage of student expuslions with and without educational services 


```{r}

 pct <- Education_stats_CLEAN %>% 
 select(disability_status_sex_and_race_ethnicity,expulsions_with_educational_services,expulsions_without_educational_services,all_expulsions) %>% 
filter(disability_status_sex_and_race_ethnicity == "All_Students") %>% 
mutate(Pct_of_all_students_with_without_explulsions = expulsions_with_educational_services/all_expulsions)  

pct=data.frame(no_students=c(66133, 45082, 111215), expulsion_type=c("With Educational Services", "Without Educational Services", "All Expulsions"))

pct_w_edu = (66133/111215)*100
pct_wo_edu = (45082/111215)*100

print(pct_w_edu)
print(pct_wo_edu)








```
#WELLS COMMENT - Please write a brief summary of what you found with this chunk and why it is significant. What does "without educational services" mean?



```{r}
#graphic 5 

pct %>% 
  ggplot() +
  geom_bar(aes(x=reorder(expulsion_type,no_students), weight=no_students), fill = "blue") +
  coord_flip() + 
  labs(
    title="differences of corpral punishment by race and gender ",
    x = "",
    y = "amount"
  )


``` 





#Most Important findings

# I originally set out to find to assess how students of different races are punished or disciplined in schools. We began by using data from NCES, or the National Center of Education statistics. The data provided statistics can be found at this link(https://nces.ed.gov/programs/digest/d18/tables/dt18_233.27.asp). Before we got a chance to assess and analyze the data, we had to clean the data which was very difficult because o the missing variables. 

#While there were many important and insightful observations we were able to derive from the data we gathered, some stuck out more than others. For example, the rate of suspension to expulsion exists for both disabled students and students in schools were quite similar even though there is a #WELLS COMMENT: NEARLY 2 MILLION DIFFERENCE. USE COMMAS TO SEPARATE AT THOUSANDS. SEE AP STYLE FOR NUMBERS#1991138 difference in the students in schools and disabled students.


#Table 3 shows us that black students get disciplined the most in schools with about 18,000 more cases of discipline than white students. This was one of the findings that stuck out to us most because it corroborated and affirmed the research and studies from organizations such as PEW research or the Brookings institutions that claim the existence of glaring disparities in the types of students who were being disciplined. #WELLS COMMENT - include a little material from pew or Brookings. Good context # While our data does not explain or show this, many studies #WELLS COMMENT - EXAMPLES OF OTHER STUDIES? # have indicated there are greater impacts on Black students later in their lives. This is why is important because our data shows 18,000 more cases of  Black students being disciplined than their white counterparts. Hispanic students faced 55,4498 suspensions compared to 18271 expulsions. Black students faced 104, 2991 suspensions compared to 33,557 expulsions. 

#Another alarming observation we noticed was the disparities in race and gender who receive corporal punishment. #WELLS COMMENT. PLEASE REVISE THIS AND PUT IN PROPER AP STYLE.#W 44,319  white males received corporal punishment compared to the 29,840 black Black Males. However,    10,183  Black females received corporal punishment compared to the 8,750 white females. This is a difference of 1433  which is a big deal. It is important to understand #WELLS COMMENT: AP STYLE. NOT CAPITALIZED# Corporal punishment uses physical force agaisnt a student to correct misbehavior. In some school districts where corporal punishment is legal, a teacher or administrator is allowed to use a  large wooden board, or paddle, to strike a child's buttocks as a form of discipline. Our findings confirm the bookings institutions claim  that "Black children are twice as likely as white children to be subject to corporal punishment" 



# More information regarding corporal punishment in schools: 
https://www.google.com/url?sa=t&rct=j&q=&esrc=s&source=web&cd=&cad=rja&uact=8&ved=2ahUKEwiF0qutiND7AhXqEFkFHWpiDi4QFnoECA8QAQ&url=https%3A%2F%2Fwww.brookings.edu%2Fblog%2Fbrown-center-chalkboard%2F2022%2F01%2F14%2Fcorporal-punishment-schools-and-race-an-update%2F&usg=AOvVaw2JpCsESNMHvWMaJa-uAQFn 


##More information regarding school to prison pipeline: 
https://www.supportiveschooldiscipline.org/school-to-prison-pipeline




