#lab_08
Sean Mussenden, Aug 16, 2022

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load janitor and the tidyverse
```{r}
library(tidyverse)
library(janitor)
```
## Load data

**Task**: Load four tables of U.S. Census data.  Each table has one row per Maryland county with 2020 population information.  Each table has these same two columns:

* geoid - a unique identifier for each county
* name - the county name

The third column in each table represents the total population in one of four demographic categories.

* maryland_population.csv - total_pop - total population 
* maryland_hispanic.csv - hispanic_pop - total Hispanic population
* maryland_white.csv - white_pop - total White alone population
* maryland_black.csv - black_pop - total Black alone population

```{r}
MD_pop <- read_csv("data/maryland_population.csv")
MD_hispanic_pop <- read_csv("data/maryland_hispanic.csv")
MD_white_pop <- read_csv("data/maryland_white.csv")
MD_black_pop <- read_csv("data/maryland_black.csv")


```

Some of the information in the tables may need to be cleaned or validated in order to produce the correct answer.  

## Questions

**Q1**  What Maryland county has the largest Hispanic population (by percentage of total population)? 
**A1** 
```{r}
#clean_names(MD_hispanic_pop)
#glimpse(MD_hispanic_pop)
MD_hispanic_pop $name<-str_replace(MD_hispanic_pop$name,"Montgomery County,Maryland", "Montgomery County, Maryland")
largest_hispanic_pop <- MD_hispanic_pop %>% 
  group_by(geoid,name) %>% 
inner_join(MD_pop,by=c("name" = "name","geoid"="geoid")) %>% 
mutate(hispanic_pct=(hispanic_pop/total_pop)*100) %>%  
  arrange(desc(hispanic_pct))

#Montgomery county 
```

**Q2**  What is the total population of people who identify as white alone in Maryland? 
**A2**  

```{r}
MD_white_pop$white_pop<-str_replace(MD_white_pop$white_pop,"13510000","13510")

MD_white_pop %>% 
as.data.frame() %>% 
clean_names() %>% 
mutate(white_pop = as.numeric(white_pop)) %>% 
summarise(
white_pop=sum(white_pop,na.rm=TRUE)
)
# total white pop is 3275048	

``` 

**Q3** What is the average Black population in all Maryland counties? 
**A3** 
```{r}
MD_black_pop[3, 3] = 243098
MD_black_pop[24, 3] = 375002

MD_black_pop %>% 
  mutate(avg_black_pop = as.numeric(black_pop)) %>% 
  summarise(avg_black_pop = mean(black_pop,na.rm=TRUE)) %>% 
  view()

#The avge Black population is 	
75130.33
``` 


  


