---
title: "R Notebook"
output: html_notebook
---
# *Notes to engineer* 
My name is Grace Kpetemey, and I worked with my partner Ben to reproduce some of the data Probliublica journalists were able to gather from police records from New York City’s Civilian Complaint Review Board. The story they were able to tell a remarkable, sad, and important. Down below you will see a link to the intresting fidnigns of the article which we will reporduce throghout this notebook and the link ot the article and data. 

#Memorable moments: Five Most Important Paragraphs 
https://www.propublica.org/article/over-a-dozen-black-and-latino-men-accused-a-cop-of-humiliating-invasive-strip-searches-the-nypd-kept-promoting-him 

#Article link: https://projects.propublica.org/nypd-ccrb/ 

```{r}
library(tidyverse)
library(janitor)
``` 


```{r}
allegations<- rio::import("allegations_202007271729.csv")
clean_names(allegations)
```
# **first article finding** 

Allegations involving McCormick: Of at least 77 allegations made against him in 26 separate CCRB complaints, 29 were unsubstantiated; five were “unfounded,” meaning investigators concluded the incident never took place; and 27 were “exonerated, 

#* *Analysis* 

This table was created to allow us to see McCormack's allegations isolated from the other incidents. This table holds various variables such as rank_incident which indicates his rank, or his level of responsibility in the NYPD. He is ranked POM or principal of operations. The complainnt_ehtnicity refers to the ethnicity of the person who complained, The board_disposition touches on the charges of the officer and what happened to the complaint. 
```{r}
nypd_mccormack <- allegations %>% 
  filter(first_name == "Christophe", last_name == "Mccormack")

nypd_mccormack %>% 
  group_by(complaint_id) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

nypd_mccormack %>% 
  filter(board_disposition == "Unsubstantiated"|board_disposition == "Exonerated") %>% 
  group_by(board_disposition) %>% 
  summarise(count = n())

``` 
# *second article finding* 

total allegations by police officer. average complaints

#*Analaysis*

The table down below indicates the number of allegations by each police officer and the average number of complaints which allows us and other readers of Propublica to get a wholistic perspective of what the NYPD and their officers were doing. 

```{r}

alleg_per_person <- allegations %>% 
select(first_name,last_name,shield_no,complaint_id) %>% 
count(first_name,last_name) %>% 
arrange(desc(n))


``` 

#Third article finding* 
allegations involving supervisors: Eighty-six of the roughly 420 officers in the department who currently hold a rank above captain — running precincts and other large commands and overseeing hundreds of officers — have tallied at 

# *Analysis*
I was able to produce a portion of the 420 officers who hold a rank above captain had at least one substantiated misconduct allegation. From the publicly released data set, we were able to identify, this table showcases officers with a rank above captain with at least one substantiated misconduct allegation.


```{r}
ranked_supervisors_allegations <- allegations %>% 
  filter(str_detect(board_disposition, "^Substantiated")) %>% 
  filter(rank_abbrev_now != "POM") %>% 
  filter(rank_abbrev_now != "POF") %>% 
  filter(rank_abbrev_now != "PO") %>% 
  filter(rank_abbrev_now != "SGT") %>% 
  filter(rank_abbrev_now != "SSA") %>% 
  filter(rank_abbrev_now != "SDS") %>% 
  filter(rank_abbrev_now != "LT") %>% 
  filter(rank_abbrev_now != "DT1") %>% 
  filter(rank_abbrev_now != "DT2") %>% 
  filter(rank_abbrev_now != "DT3") %>% 
  filter(rank_abbrev_now != "DTS") %>% 
  filter(rank_abbrev_now != "CPT") %>% 
  filter(rank_abbrev_now != "LCD") %>% 
  filter(rank_abbrev_now != "LSA") 

ranked_supervisors_allegations%>% 
  group_by(first_name, last_name, unique_mos_id) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total))


```
# fourth article finding
Misconduct allegations by race of victim

#*analysis* 
This table indicates the misconduct allegations and the race of the individuals who made the complaints. This is particularly interesting because it allows you to see the very clear pattern and racial theme among the complainants. 

```{r}
misc_race_victim <-allegations %>% 
  select(fado_type, complainant_ethnicity, allegation) 


``` 


